\thetitle

### 2009-02-19 paper/paper.tex
ispell -b for british spelling
darcs-hash:20090219162103-d98fc-037b1406047bacd85202f9131518a43a1c1c0619.gz

The key to giving a proper formal treatment of both these aspects lies in using an appropriate representation of the algorithm, capable of capturing its computational behavior---specifically, its use of comparisons---in a way suitable for subsequent formal reasoning. 
The appproach we take is to consider such operation-counting as a \emph{side effect}, and to use the general framework of \emph{monads} for representing side-effecting computation in pure functional languages. 
Accordingly we use a shallow embedding, in which the algorithm, here Quicksort, is written as a monadically expressed functional program in Coq. This definition is then instantiated with refinements of operation-counting monads to make the comparison count observable.

@@

Section \ref{finishing} finishes the proof by proving the lemma about the expected comparison count for individual input list elements. Since this is the part where the original proof omits the most detail, and makes the assumption regarding duplicate elements, and where we really have to reason in detail about the behavior of the algorithm, it is by far the most involved part of the formalisation.

@@

Note that the deterministic Quicksort definition in section \ref{embed} could also have been implemented with a |partition| pass instead, which might well have made the worst-case proof even simpler. We chose not to do this, in order to emphasize that the properties the average-case proof demands of the algorithm rule out the na{\"\i}ve but familiar implementation using |filter| passes.

@@

>>>>

The key to giving a proper formal treatment of both these aspects lies in using an appropriate representation of the algorithm, capable of capturing its computational behaviour---specifically, its use of comparisons---in a way suitable for subsequent formal reasoning. 
The approach we take is to consider such operation-counting as a \emph{side effect}, and to use the general framework of \emph{monads} for representing side-effecting computation in pure functional languages. 
Accordingly we use a shallow embedding, in which the algorithm, here Quicksort, is written as a monadically expressed functional program in Coq. This definition is then instantiated with refinements of operation-counting monads to make the comparison count observable.

@@

Section \ref{finishing} finishes the proof by proving the lemma about the expected comparison count for individual input list elements. Since this is the part where the original proof omits the most detail, and makes the assumption regarding duplicate elements, and where we really have to reason in detail about the behaviour of the algorithm, it is by far the most involved part of the formalisation.

@@

Note that the deterministic Quicksort definition in section \ref{embed} could also have been implemented with a |partition| pass instead, which might well have made the worst-case proof even simpler. We chose not to do this, in order to emphasise that the properties the average-case proof demands of the algorithm rule out the na{\"\i}ve but familiar implementation using |filter| passes.

@@

% LocalWords:  Nijmegen GL Quicksort Coq todo monadically Hurd Audebaud Paulin
% LocalWords:  Mohring na ve Coq's pCIC ret Haskell's liftM forall filterM bool
% LocalWords:  parameterizing le gt negb Fixpoint qs Quicksort's customized MMT
% LocalWords:  specialized elementwise idealized monoid fcmp ab snd mult fst SP
% LocalWords:  nat NatAddMonoid IdMonad Eval infixbind sqrd flt Eq cmp const ne
% LocalWords:  upp avg ary expec NDP MonoidMonadTrans NeTreeMonad bigO INR ceil
% LocalWords:  IJ ecc li fr NotIn generalizing ca cb foo parameterized
